---
title: "Tract to CCA"
output: html_document
date: "2023-02-10"
---
# for installing packages in this code (you only need to do this once)
```{r}
install.packages("dplyr")
install.packages("tidyverse")
install.packages("tidycensus")
install.packages("writexl")
install.packages("Rtools")
```
# for loading packages for all future uses 
```{r}
devtools::install_github("CMAP-REPOS/cmapgeo")
library(dplyr)
library(tidyverse)
library(tidycensus)
library(cmapgeo)
library(tidyverse)
library(writexl)
```
# View the tracts with population split between multiple CCAs, filter for greatest value among duplicates, and export as excel table.
# You will create two outputs - a csv and xlsx. Use the csv to view in R and use the xlsx (access externally in your folder, outside of RStudio) for external data analysis.
```{r}
filter(xwalk_tract2cca, pop_pct < 1)
Pop_geocode <-xwalk_tract2cca %>% select('geoid_tract', 'cca_num', 'pop_pct') %>% group_by(geoid_tract)%>% filter(pop_pct == max(pop_pct))

write.csv(Pop_geocode, file = "pop_geocode.csv", quote = FALSE, row.names = FALSE)
write_xlsx(Pop_geocode, "pop_geocode.xlsx", col_names = TRUE, format_headers = TRUE)

```
# Repeat above but for housing units, households, and employment. 
#View the tracts with housing units plit between multiple CCAs, filter for greatest value among duplicates, and export as excel table.
# You will create two outputs - a csv and xlsx. Use the csv to view in R and use the xlsx (access externally in your folder, outside of RStudio) for external data analysis.
```{r}
filter(xwalk_tract2cca, hu_pct < 1)
Hu_geocode <-xwalk_tract2cca %>% select('geoid_tract', 'cca_num', 'hu_pct') %>% group_by(geoid_tract)%>% filter(hu_pct == max(hu_pct))

write.csv(Hu_geocode, file = "Hu_geocode.csv", quote = FALSE, row.names = FALSE)
write_xlsx(Hu_geocode, "Hu_geocode.xlsx", col_names = TRUE, format_headers = TRUE)

```
```{r}
filter(xwalk_tract2cca, hh_pct < 1)
Hh_geocode <-xwalk_tract2cca %>% select('geoid_tract', 'cca_num', 'hh_pct') %>% group_by(geoid_tract)%>% filter(hh_pct == max(hh_pct))

write.csv(Hh_geocode, file = "Hh_geocode.csv", quote = FALSE, row.names = FALSE)
write_xlsx(Hh_geocode, "Hh_geocode.xlsx", col_names = TRUE, format_headers = TRUE)

```
```{r}
filter(xwalk_tract2cca, emp_pct < 1)
Emp_geocode <-xwalk_tract2cca %>% select('geoid_tract', 'cca_num', 'emp_pct') %>% group_by(geoid_tract)%>% filter(emp_pct == max(emp_pct))

write.csv(Emp_geocode, file = "Emp_geocode.csv", quote = FALSE, row.names = FALSE)
write_xlsx(Emp_geocode, "Emp_geocode.xlsx", col_names = TRUE, format_headers = TRUE)

```

# Estimate CCA-level housing vacancy rates from block group-level Census data
```{r}
xwalk_tract2cca
df_tract <- tidycensus::get_decennial(
  geography = "tract", variables = c("H1_001N", "H1_003N"),
  year = 2020, state = "IL", county = c("031", "043"), output = "wide"
) %>%
  suppressMessages() %>%  
  select(geoid_tract = GEOID, hu_tot = H1_001N, hu_vac = H1_003N)

```
```{r}
df_tract <- xwalk_tract2cca %>%
  left_join(df_tract, by = "geoid_tract") %>%
  mutate(hu_tot = hu_tot * hu_pct,
         hu_vac = hu_vac * hu_pct) %>%
  group_by(geoid_tract) %>%
  summarize_at(vars(hu_tot, hu_vac), sum) %>%
  mutate(vac_rate = hu_vac / hu_tot)

df_tract
```

# Join to cca_sf for mapping
```{r}
library(ggplot2)
tract_sf %>%
  left_join(df_tract, by = "geoid_tract") %>%
  ggplot() +
    geom_sf(aes(fill = vac_rate), lwd = 0.1) +
    scale_fill_viridis_c(direction = -1) +
    theme_void()
```

